### 2018-04-24 README.Rmd
initial commit

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
library(deeptriangle)
library(tidyverse)
library(keras)
```
# DeepTriangle

This is the companion repository to the DeepTriangle paper.

## Reproducing experiments

To get started, either clone the repo and build the R package, or install with

``` r
devtools::install_github("kevinykuo/deeptriangle)
```

You will also need the [insurance](https://github.com/kevinykuo/insurance) package, which can be installed with

```r
devtools::install_github("kevinykuo/insurance")
```

The following lines of code will instantiate and fit the model for each line of business then combine the forecasts in a data frame:

```{r, message = FALSE, cache = TRUE}
library(deeptriangle)
library(tidyverse)
library(keras)

# set seed for reproducibility
use_session_with_seed(2018)

data <- dt_data_prep(insurance::schedule_p, dt_group_codes)

lobs <- c("workers_compensation", "commercial_auto",
          "private_passenger_auto", "other_liability")

predictions <- lobs %>%
  map(
    function(x) {
      # clear session and instantiate model
      k_clear_session()
      model <- dt_model()

      c(training_data, validation_data, full_training_data) %<-%
        dt_train_validation_split(data[[x]])

      message("Training - ", x)

      # determine number of epochs
      epochs_to_train <- dt_optimize_epochs(
        model, training_data, validation_data
      )

      # clear session and reinstantiate model
      k_clear_session()
      model <- dt_model()

      # fit model to all training data
      history <- model %>%
        fit(x = full_training_data$x,
            y = full_training_data$y,
            batch_size = 128,
            epochs = epochs_to_train,
            verbose = 0)
      dt_compute_predictions(model, data[[x]])
    }) %>%
  bind_rows()
```

We can then compute performance metrics...

```{r}
model_results <- dt_compute_metrics(predictions) %>%
  bind_rows(stochastic_model_results) %>%
  gather(metric, value, mape, rmspe)
```

and tabulate the results:

```{r}
dt_tabulate_metrics(model_results, metric = "mape") %>%
  knitr::kable(booktabs = "T", digits = 3)
```

```{r}
dt_tabulate_metrics(model_results, metric = "rmspe") %>%
  knitr::kable(booktabs = "T", digits = 3)
```

To create actual vs. predicted plots, use the `dt_plot_predictions()` function.

```{r, out.width = "60%"}
dt_plot_predictions(predictions, "337", "workers_compensation",
                    "paid_loss")
```

```{r, out.width = "60%"}
dt_plot_predictions(predictions, "337", "workers_compensation",
                    "claims_outstanding")
```